Evidence of the necessity of isotopic precise calculation in fuel cycle calculation

### 2016-07-01 paper.tex
final edits

All the calculation of this study have been performed using the Cyclus fuel cycle
tool\cite{CYCLUS}. The facility models are from the Cycamore package, developed by the
Cyclus team, and containing a Pu-equivalent fuel fabrication model as well as
a simple fixed mixing ratio fuel fabrication model. A new neural network model has
also been contributed to the Cyclus environment using the cyCLASS tool\cite{cyCLASS}, 
allowing the use of the fuel fabrication and depletion models from the CLASS tool
\cite{CLASS} in Cyclus.

@@

This work compares the plutonium fraction required to build fresh MOX fuel
using three different stream mixing models in fuel fabrication: ``fixed mixing
ratio'', ``plutonium equivalent theory'', ``neural network'' \cite{Leniau2015125}.
The first part of the study in focuses on the effect of decay on the isotopic
composition of the reprocessed plutonium and its impact on fuel fabrication
depending on the chosen stream mixing model.  The second part of the study,
compares the impact on fresh fuel compositions when used fuel is based on a
depletion calculation instead of a fixed recipe, using the neural network fuel
fabrication model, also including the impact of decay.

@@

\subsection{Fabrication model}
Three different ways to model the fuel fabrication have been used: ``fix mixing
ratio'', ``Pu-equivalent'' and ``neural network''.
The fix mixing ratio is based on a user-defined ratio for the J1 and E3 stream
mixing. The respective mixing ratio used for J1 and E3 stream are 
$0.8357$ and $0.1643$.
The ``Pu-equivalent'' model finds the mixing ratio that results in an initial
reactivity that matches that of the requested MOX fuel recipe.  This ratio may
change as the composition of the incoming streams changes.  As the requested
fuel composition corresponds to the fuel composition built with the ''fix mixing
ratio`` model, the results of those 2 models are expected to be similar.  The
``neural network'' model attempts to achieve a given reactivity at the  end of
cycle reactivity of the whole core for a multi-batch core ($50~$GWd/t in this
study). The $k_{infty}$ used in this study is 1.034, assuming the reactivity of
the full core is the mean reactivity of the batches. The neural network in this
model has been pre-trained on several different depletion calculations, allowing
it to predict the correct plutonium fraction required, depending on the
composition of that plutonium.

Because the neural network model allows only to mix a plutonium stream into a
uranium stream, slight modification have been made on the cycle: in both ``J1''
and ``E3'' streams uranium and plutonium are separated, the uranium and
plutonium sub-streams are mix together to build one stream of pure uranium and
one stream of pure plutonium. That is why, when comparing the different fuel
fabrication models only the total plutonium enrichment are compared.

Because of these models have not necessary been tuned on exactly the same
PWR configuration and parameters, this analysis focuses on the differences
in the results due to the change in model rather than the results themselves.

The second part of the study compares the amount of plutonium require to build
the PWR-MOX fuel (using the neural network modeling for fuel fabrication) for
the recipe based output composition versus a composition calculated by
depletion.  To calculate the depletion for each fuel batch loaded in the PWR
reactor, another neural network is used to predict the macroscopic cross section
of the fuel \cite{Leniau2015125} (only fission, capture and n,2n reactions are
considered). The depletion calculation is done renormalizing the neutron flux to
maintain a constant power in the reactor.

@@

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.48\textwidth]{nodecay_pu_contribution.png}
  \caption{Evolution of the fraction of plutonium in the fresh PWR-MOX fuel
  using the 3 predicted models: ``fix mixing ratio'' (``fix'', blue), ``Pu-equivalent''
  (``Pu-eq'', green) and ``Neural Network'' (``NN'', red), without considering decay.}
  \label{fig:nod}
\end{figure}

@@

\begin{figure}[h!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.48\textwidth]{decay_pu_contribution.png}
  \caption{Evolution of the fraction of plutonium in the fresh PWR-MOX fuel
    using the 3 prediction models: ``fix mixing ratio'' (``fix'', blue),
    ``Pu-equivalent'' (``Pu-eq'', green) and ``Neural Network'' (``NN'', red),
    with (solid line) and without (dash line) decay enable.} 
  \label{fig:d}
\end{figure}

@@

A last calculation, on decay have been performed using the ``Pu-equivalent''
fabrication model, increasing by a factor 2, 5 and 10 the amount of plutonium in
the startup inventory. This mimics different degrees of a plutonium accumulation,
and thus different impacts from the decay of this plutonium.
This calculation results in a change in the mixing ratio required to build the
PWR-MOX fuel of 10 to 30\%. This kind of plutonium accumulation can
unexpectedly occur in a real fuel cycle and might need to be considered in a
complete fuel cycle transition study.

@@

Four simulations were carried out representing variations across two problem
dimensions.  One variation compares how reactor models determine the recipe of
used fuel.  The standard recipe reactor model produces used fuel of a fixed
composition, regardless of the composition of the incoming fuel.  The
alternative is a neural network depletion model provided by the cyCLASS
package.  This model determines the composition of the used fuel depending on
its initial composition and discharge burnup.  Both of these models were used
without decay and with decay.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.48\textwidth]{irradiation_pu_contribution.png}
  \caption{Evolution of the fraction of plutonium in the fresh PWR-MOX fuel
    using the ``Neural Network'' prediction models, using recipe based spent
    fuel composition (red) and recalculated spent fuel composition (orange),
    with (solid line) and without (dash line) decay enable.}
  \label{fig:depletion}
\end{figure}

@@

This is caused by the slight change in the composition in the available
plutonium (see Fig. \ref{fig:depletioncompo}): in the case of the depletion
calculation the available plutonium contain $68\%$ of $^{239}$Pu versus $64\%$
in the recipe based case, and also less $^{242}$Pu and $^{241}$Am, respectively
$4.8\%$ and $5.4\%$ in the depletion case versus $7.3\%$ and $6.7\%$. The
amounts of the other nuclides are very close.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.48\textwidth]{MOX_pu_composition.png}
  \caption{Evolution of the composition of the plutonium in the fresh PWR-MOX fuel
  using the ``Neural Network'' prediction models, with decay enable using recipe
  based spent fuel composition (plain line) and recalculated spent fuel
  composition (dash line).}
  \label{fig:depletioncompo}
\end{figure}

Those slight change in the output composition, lead to an overall decrease in
the amount of $^{241}$Am in the fresh PWR-MOX fuel and an increase of the
$^{239}$Pu, allowing to build fuel with a lower plutonium enrichment.

>>>>

All the calculation of this study have been performed using the Cyclus fuel cycle
simulator framework\cite{CYCLUS}. The facility models are from the Cycamore package, developed by the
Cyclus team, and containing a Pu-equivalent fuel fabrication model as well as
a simple fixed mixing ratio fuel fabrication model. A new neural network model has
also been contributed to the Cyclus environment using the cyCLASS tool\cite{cyCLASS}, 
allowing the use of the fuel fabrication and depletion models from the CLASS tool
\cite{CLASS} in Cyclus.

@@

This work compares the plutonium fraction required to build fresh MOX fuel
using three different stream mixing models in fuel fabrication: ``fixed mixing
ratio'', ``plutonium equivalent theory'', ``neural network'' \cite{Leniau2015125}.
The first part of the study focuses on the effect of decay on the isotopic
composition of the reprocessed plutonium and its impact on fuel fabrication
depending on the chosen stream mixing model.  The second part of the study,
compares the impact on fresh fuel compositions when used fuel is based on a
depletion calculation instead of a fixed recipe, using the neural network fuel
fabrication model, also including the impact of decay.

@@

\subsection{Fabrication model}
Three different ways to model the fuel fabrication have been used: ``fix mixing
ratio'', ``Pu-equivalent'' and ``neural network''.
The fixed mixing ratio is based on a user-defined ratio for the J1 and E3 stream
mixing, with mixing ratio of $0.8357$ and $0.1643$, respectively.
The ``Pu-equivalent'' model finds the mixing ratio that results in an initial
reactivity that matches that of the requested MOX fuel recipe.  This ratio may
change as the composition of the incoming streams changes.  As the requested
fuel composition corresponds to the fuel composition built with the ``fix mixing
ratio'' model, the results of those 2 models are expected to be similar.  The
``neural network'' model attempts to achieve a given whole core end of cycle 
reactivity for a multi-batch core with specified discharge burnup ($50~$GWd/t in this
study). The $k_{infty}$ used in this study is 1.034, assuming the reactivity of
the full core is the mean reactivity of the batches. The neural network in this
model has been pre-trained on several different depletion calculations, allowing
it to predict the correct plutonium fraction required, depending on the
composition of that plutonium.

Because the neural network model allows only mixing of a plutonium stream into a
uranium stream, slight modification have been made on the cycle: in both ``J1''
and ``E3'' streams uranium and plutonium are separated, the uranium and
plutonium sub-streams are mix together to build one stream of pure uranium and
one stream of pure plutonium. That is why, when comparing the different fuel
fabrication models only the total plutonium enrichment are compared.

Because these models have not necessarily been tuned on exactly the same
PWR configuration and parameters, this analysis focuses on the differences
in the results due to the change in model rather than the results themselves.

The second part of the study compares the amount of plutonium required to build
the PWR-MOX fuel (using the neural network modeling for fuel fabrication) for
the recipe based output composition versus a composition calculated by
depletion.  To calculate the depletion for each fuel batch loaded in the PWR
reactor, another neural network is used to predict the macroscopic cross section
of the fuel \cite{Leniau2015125} (only fission, capture and n,2n reactions are
considered). The depletion calculation is done renormalizing the neutron flux to
maintain a constant power in the reactor.

@@

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.48\textwidth]{nodecay_pu_contribution.png}
  \caption{Evolution of the fraction of plutonium in the fresh PWR-MOX fuel
  using the 3 predicted models: ``fix mixing ratio'' (``fix'', blue), ``Pu-equivalent''
  (``Pu-eq'', green) and ``neural network'' (``NN'', red), without considering decay.}
  \label{fig:nod}
\end{figure}

@@

\begin{figure}[h!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.48\textwidth]{decay_pu_contribution.png}
  \caption{Evolution of the fraction of plutonium in the fresh PWR-MOX fuel
    using the 3 prediction models: ``fix mixing ratio'' (``fix'', blue),
    ``Pu-equivalent'' (``Pu-eq'', green) and ``neural network'' (``NN'', red),
    with (solid line) and without (dash line) decay enable.} 
  \label{fig:d}
\end{figure}

@@

Finally, a set of calculations with decay have been performed using the ``Pu-equivalent''
fabrication model, increasing by a factor 2, 5 and 10 the amount of plutonium in
the startup inventory. This mimics different degrees of a plutonium accumulation,
and thus different impacts from the decay of this plutonium.
This calculation results in a change in the mixing ratio required to build the
PWR-MOX fuel of 10 to 30\%. This kind of plutonium accumulation can
unexpectedly occur in a real fuel cycle and might need to be considered in a
complete fuel cycle transition study.

@@

Four simulations were carried out representing variations across two problem
dimensions.  One variation compares how reactor models determine the recipe of
used fuel.  The standard recipe reactor model produces used fuel of a fixed
composition, regardless of the composition of the incoming fuel.  In this case, the
alternative is a neural network depletion model provided by the cyCLASS
package.  This model determines the macroscopic cross sections of the fuel as a
function of burnup depending on its initial composition and discharge burnup. It then
performs a simple calculation to allow the composition to evolve. Both of these models were used
without decay and with decay.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.48\textwidth]{irradiation_pu_contribution.png}
  \caption{Evolution of the fraction of plutonium in the fresh PWR-MOX fuel
    using the ``neural network'' prediction models, using recipe based spent
    fuel composition (red) and recalculated spent fuel composition (orange),
    with (solid line) and without (dash line) decay enable.}
  \label{fig:depletion}
\end{figure}

@@

This is caused by the slight change in the composition in the available
plutonium (see Fig. \ref{fig:depletioncompo}): in the case of the depletion
calculation the available plutonium contain $68\%$ of $^{239}$Pu versus $64\%$
in the recipe based case, and also less $^{242}$Pu and $^{241}$Am: respectively
$4.8\%$ and $5.4\%$ in the depletion case versus $7.3\%$ and $6.7\%$. The
amounts of the other nuclides are very close.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.48\textwidth]{MOX_pu_composition.png}
  \caption{Evolution of the composition of the plutonium in the fresh PWR-MOX fuel
  using the ``neural network'' prediction models, with decay enable using recipe
  based spent fuel composition (plain line) and recalculated spent fuel
  composition (dash line).}
  \label{fig:depletioncompo}
\end{figure}

Those slight change in the output composition, lead to an overall decrease in
the amount of $^{241}$Am in the fresh PWR-MOX fuel and an increase of the
$^{239}$Pu, allowing fuel to be built with a lower plutonium enrichment.
